## **R Libraries: ggplot2**

The R library `ggplot2` is a super useful tool we can use to create visualizations and explore our data.

```{bash}
install.packages("tidyverse")
library("tidyverse")
```

More things we can do to get more information on R and what we're doing:

```{bash}
help()
```

We'll need a few more packages for our exploration usage, too. We can install them all at once by using the `c()` function, which creates a list. We'll open these packages with the `library()` function later.

```{bash}
install.packages(c("nycflights13", "gapminder", "Lahman"))
```

Built in to our tidyverse are several data sets that we can play around with. In the below code block, we create a variable `myplot` and save the `mpg` database to it.

```{r}
myplot <- mpg
myplot
```

**Something you should be aware of and have open (perhaps on a second monitor, or you can print them out) is the ggplot2 cheat sheet!** <https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf>

```{r}
ggplot(data = mpg) +
  geom_point(mapping =aes(x=displ, y=hwy))

```

```{r}
mtcars
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y=hwy, color = class))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y=hwy, color = displ < 5))
view(mpg)

```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y =hwy , color = rgb(0.7,1,0.3))) +
  facet_wrap(~class, nrow = 3)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ , y =hwy , color = class))+
  facet_wrap(~class, nrow = 3)
```

```{r}
ggplot(data=mpg) +
  geom_point(mapping =aes(x = displ , y = hwy, color = drv)) + 
  facet_grid(drv ~ .)
```

```{r}
#ggplot(data = mpg) +
   #geom_smooth(mapping = aes(x=displ, y =hwy , linetype = drv)) +
   #geom_point(mapping = aes(x= displ , y = hwy , color =drv))

ggplot(data = mpg , mapping = aes( x= displ, y= hwy , linetype = drv)) +
  geom_point(mapping = aes(color =class)) +
  geom_smooth( se =FALSE)
  
```

Whoa, where'd count come from? **Count is not a variable or attribute of the variable diamonds!** Count comes automatically with a lot of graphs as an attribute. You can verify what I'm saying with `?diamonds` in the console. It shows up this is a discrete data set, rather than the contionius data set the we have been using

Any attribute of a data set that is algorithmically calculated is called a *stat*, which is short for a statistical transformation. Many of the **geom** functions have stats built in, and many stats display geoms. For example, the above code block used a geom, but this one uses a stat, and results in an identical chart:

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

Let's add color here!

```{r}
ggplot(data= diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

What if we add to a different variable than cut?

```{r}
pie= ggplot(data = diamonds) +
  geom_bar(mapping =aes(x = cut, fill = clarity),
           position = "fill"
           ) +
  theme(aspect.ratio = 1) +
  labs(x = NULL , y = NULL)

pie + coord_polar()
```

**For your consideration...** Is the above a good graph?

```{r}
#alt- becomes <- (option- on mac)

pie= ggplot(data = diamonds) +
  geom_bar(mapping =aes(x = cut, fill = clarity),
           position = "dodge"
           ) +
  theme(aspect.ratio = 1) +
  labs(x = NULL , y = NULL)

pie + coord_polar()
```

```{r}

```

## Homework

```{r}
# Find a builtin data set for students to explore; use their cheat sheet from stats to label continuous and discrete data; Use graphs to tell me things about 3 data sets 
#Analyze that data set, create graphs, find a hypothesis

#View(nycflights13::flights)
#View(diamonds)

ggplot(data = iris) +
  geom_bar(mapping = aes(x= Species , fill= Species))

# Iris has three Species type and each are 50 in number as it is displayed on the bar chart




```

```{r}
ggplot(data = nycflights13::flights) +
  geom_point(mapping = aes(x = dep_time, y=sched_dep_time, color = dep_time))
 
# The scatter plot shows the relationship between the departure time and schedule departure time on the nycflights data frame
```




```{r}
ggplot(data = nycflights13::flights) +
```

```{r}

#View(trees)
ggplot(data = trees) +
  geom_point(mapping = aes( x= Girth, y= Volume, color = Girth ))

# trees is a data frame that has 31 observations on 3 variables that provides measurements of diameter, height and volume for
#Black Cherry Trees. Using ggplot and scatter plots it is shown the relationship between the Girth and the voloume of a tree


```
```

```         
```

`
```

```         
```
